### 2020-07-23 README.rmd
docs(readme,home): update branch descriptions from master to main

---
output: github_document
---

<!-- README.md is generated from README.rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# ggalluvial

[![Travis](https://travis-ci.org/corybrunson/ggalluvial.svg?branch=main)](https://travis-ci.org/corybrunson/ggalluvial)
[![CRAN](http://www.r-pkg.org/badges/version/ggalluvial)](https://cran.r-project.org/package=ggalluvial)

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3836748.svg)](https://doi.org/10.5281/zenodo.3836748)
[![JOSS](https://joss.theoj.org/papers/10.21105/joss.02017/status.svg)](https://doi.org/10.21105/joss.02017)

This is a [**ggplot2** extension](https://exts.ggplot2.tidyverse.org/) for alluvial plots.

## Design

The alluvial plots implemented here can be used to visualize frequency distributions over time or frequency tables involving several categorical variables. The design is inspired by the [**alluvial**](https://github.com/mbojan/alluvial) package, but the **ggplot2** framework induced several conspicuous differences:

- **alluvial** understands a variety of inputs (vectors, lists, data frames), whereas **ggalluvial** requires a single data frame;
- **alluvial** uses each variable of these inputs as a dimension of the data, whereas **ggalluvial** requires the user to specify the dimensions, either as separate aesthetics or as [key-value pairs](http://tidyr.tidyverse.org/);
- **alluvial** produces both the *alluvia*, which link cohorts across multiple dimensions, and (what are here called) the *strata*, which partition the data along each dimension, in a single function; whereas **ggalluvial** relies on separate layers (stats and geoms) to produce strata, alluvia, and alluvial segments called *lodes* and *flows*.

Additionally, **ggalluvial** arranges these layers vertically without gaps, so that the secondary plotting axis indicates the cumulative values of the strata at each dimension.

## Installation

The latest stable release can be installed from CRAN:
```{r eval=FALSE}
install.packages("ggalluvial")
```

The [`cran` branch](https://github.com/corybrunson/ggalluvial/tree/cran) will contain the version most recently submitted to [CRAN](https://cran.r-project.org/package=ggalluvial). It is duplicated in the `master` branch, from which source the [website](https://corybrunson.github.io/ggalluvial/) is built.

The development version can be installed from the (default) `main` branch on GitHub:
```{r eval=FALSE}
remotes::install_github("corybrunson/ggalluvial@main", build_vignettes = TRUE)
```
Note that, in order to build the vignettes, the imported packages **alluvial**, **ggfittext**, and **ggrepel** must be installed. To skip this step, leave `build_vignettes` unspecified or set it to `FALSE`.

The
[`optimization` branch](https://github.com/corybrunson/ggalluvial/tree/optimization) contains a development version with experimental functions to reduce the number or area of alluvial overlaps (see issue [#6](https://github.com/corybrunson/ggalluvial/issues/6)). Install it as follows:
```{r eval=FALSE}
remotes::install_github("corybrunson/ggalluvial", ref = "optimization")
```
Note, however, that this branch has not kept pace with the `main` branch or with recent upgrades on CRAN.

## Usage

```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
devtools::load_all()
```

### Example

Here is how to generate an alluvial plot representation of the multi-dimensional categorical dataset of passengers on the Titanic:

```{r eval=TRUE}
titanic_wide <- data.frame(Titanic)
head(titanic_wide)
ggplot(data = titanic_wide,
       aes(axis1 = Class, axis2 = Sex, axis3 = Age,
           y = Freq)) +
  scale_x_discrete(limits = c("Class", "Sex", "Age"), expand = c(.2, .05)) +
  xlab("Demographic") +
  geom_alluvium(aes(fill = Survived)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  ggtitle("passengers on the maiden voyage of the Titanic",
          "stratified by demographics and survival")
```

The data is in "wide" format, but **ggalluvial** also recognizes data in "long" format and can convert between the two:

```{r eval=TRUE}
titanic_long <- to_lodes_form(data.frame(Titanic),
                              key = "Demographic",
                              axes = 1:3)
head(titanic_long)
ggplot(data = titanic_long,
       aes(x = Demographic, stratum = stratum, alluvium = alluvium,
           y = Freq, label = stratum)) +
  geom_alluvium(aes(fill = Survived)) +
  geom_stratum() + geom_text(stat = "stratum") +
  theme_minimal() +
  ggtitle("passengers on the maiden voyage of the Titanic",
          "stratified by demographics and survival")
```

### Documentation

For detailed discussion of the data formats recognized by **ggalluvial** and several examples that illustrate its flexibility and limitations, read the technical vignette:

```{r eval=FALSE}
vignette(topic = "ggalluvial", package = "ggalluvial")
```

The documentation contains several examples; use `help()` to call forth examples of any layer (`stat_*` or `geom_*`) or of the conversion functions (`to_*_form`).

## Acknowledgments

### Resources

Development of this package benefitted from the use of equipment and the support of colleagues at [UConn Health](https://health.uconn.edu/).

### Contribute

Contributions in any form are more than welcome! Pretty much every fix and feature of this package derives from a problem or question posed by someone with datasets or design goals i hadn't anticipated. See the [CONTRIBUTING](https://github.com/corybrunson/ggalluvial/blob/main/CONTRIBUTING.md) file for guidance, and please respect the [Code of Conduct](https://github.com/corybrunson/ggalluvial/blob/main/CODE_OF_CONDUCT.md).

### Cite

If you use **ggalluvial**-generated figures in publication, i'd be grateful to hear about it! You can also cite the package according to `citation("ggalluvial")`.